---
title: "macroStan"
output:
  github_document: default
---

```{r setup, echo=FALSE, cache = FALSE}
knitr::opts_chunk$set(echo = TRUE)
wd = "."
knitr::read_chunk(file.path(wd, "R/macro_ncp.R"), labels = "npc_macro", from = 2) # skip roxygen tags
knitr::read_chunk(file.path(wd, "R/macro_hs.R"), labels = "hs_macro", from = 8) # skip roxygen tags


input_file = file.path(wd, "inst/stan_scaffolds/macro_ncp.stan")
input_file_hs = file.path(wd, "inst/stan_scaffolds/macro_hs.stan")
knitr::read_chunk(input_file, label = "npc_scaffold")
knitr::read_chunk(input_file_hs, label = "hs_scaffold")


library(macroStan)

```

R package for defining text macros in Stan for flexible code reuse.

macroStan provides a method to define reuseable, parameterized chunks of code and insert them into the appropriate locations of a stan file.  The package was inspired by (and uses) the `glue` package.

## Example: Non-centered parameterization

As a simple example, we'll define a macro for the non-centered parameterization of a hierarchical parameter.  

```{r npc_macro, echo=TRUE, eval = FALSE}
```

This defines a macro with the arguments `name`, `mu`, `sd`, and `size`. The macro contains a number of sections (`coef`, `parms`, `tp1`, `prior`), which contain snippets of stan code.  The macro's arguments are referenced in the sections with double curly braces. 
Let's use this macro to setup a hierarchical intercept with specific parameters:

```{r echo=TRUE, eval = TRUE}
alpha_ncp = stan_macro_ncp(name = "alpha", mu = "mu", sd = "tau", size = "N_groups")
```

Now the parameters snippet (`alpha_ncp$parm`) is:

```{r, echo = FALSE} 
cat(alpha_ncp$parm)
``` 

Once a macro has been defined, a stan model can be written to use it

```{stan npc_scaffold, echo = TRUE, output.var = "junk", eval = FALSE}
```

There are two ways of inserting macros into a stan file: whole line replacement, which begins with a `$`, and in-line replacement, which uses the same `{{ }}` as the macro definitions.  To put them together, use `parse_stan_macros`.

```{r, echo = TRUE, eval = TRUE, results = 'hide', message=FALSE} 
output_file = tempfile(fileext = ".stan")
# input_file corresponds to the above stan model
parse_stan_macros(input = input_file, out_file = output_file,
                  alpha_ncp = alpha_ncp)
```

Which produces the following stan file:
```{r, echo = FALSE, eval = TRUE}
knitr::read_chunk(output_file, label = "ncp_out")

```
```{stan ncp_out, echo = TRUE, output.var = "junk2", eval = FALSE}
```

## Example 2: Horseshoe priors

Stan macros are intended for more complex code.  As an example, here is an adapted version of the horseshoe prior used in `brms`.

```{r hs_macro, echo = TRUE, eval = TRUE}
```

`input_file_hs:`
```{stan hs_scaffold, echo = TRUE, output.var = "junk3", eval = FALSE}
```

```{r, echo = TRUE, eval = TRUE, results = 'hide', message=FALSE}
output_file_hs = tempfile(fileext = ".stan")
parse_stan_macros(input_file_hs, output_file_hs, 
                  hs_betas = stan_macro_horseshoe("1", "D"))
```

Output:
```{r, echo = FALSE, eval = TRUE}
knitr::read_chunk(output_file_hs, label = "hs_out")

```
```{stan hs_out, echo = TRUE, output.var = "junk4", eval = FALSE}
```
